---
title: "Working with spatial data in R"
subtitle: "https://github.com/ecodatasci-tlv/spatial"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["default", "custom-fonts.css"]
    nature:
      highlightStyle: github
      highlightLines: true
      ratio: '16:9'
      beforeInit: "https://platform.twitter.com/widgets.js"
---

```{r setup, include=FALSE}
library(knitr)
library(tidyverse)
conflicted::conflict_prefer("filter", "dplyr")

options(htmltools.dir.version = FALSE,
        tibble.width = 70,
        ggplot2.continuous.colour = "viridis",
        ggplot2.continuous.fill = "viridis")
opts_chunk$set(
  fig.width = 6.5,
  fig.height = 4.5,
  fig.align = "center",
  cache = TRUE
)
theme_set(theme_minimal() +
            theme(text = element_text(size = 20))) 
```

# Topics we'll cover today:

* Types of spatial data
* Importing/Exporting spatial data
* Projections, extents, and units
* Comparison of different GIS tools out there
* Spatial data manipulation
* Intro to spatial analysis
* Visualizations (if we have time)

---
class: inverse, middle, center

#Data types




*https://www.gislounge.com/geodatabases-explored-vector-and-raster-data/
  
---
class: inverse, middle, center

# Vector data
---

### Polygons 
Two dimentional and are used for measurment of area and paramiters ofa geographical feature. 

---

### Lines
One dimentional and are used to only to measure length
---

### Points
Used to represent nonadjucent features 

---

IMPORTANT: In species distribution polygons are usually an over representation of the real distribtion, while point data are an under representation.
---
class: inverse, middle, center

# Raster data

Cell-based data, represents surfaces. Each cell 
---

### Continuous 
Example: temperature, elevation, etc.

### Discrete
Example: population density, species richenss, etc.


---

Rasters have three types of datasets: 

### thematic
Used to represent discrete data, such as soil type or land use
ADD A PICTURE
---
### spectral
Aerial or setalite imagery
ADD A PICTURE
---
### Surface
Represents continuous changes across a landscape (for example elevation)
ADD A PICTURE

---


# Importing Spatial Data

Now that we've learned about different types of spatial data, let's load some. We'll use the rgdal package for this:

```{r rgdal, echo = TRUE}
library(rgdal)

pkmng_countries <- readOGR(dsn = "Data",
                           layer = "pkmng_polygons")
head(pkmng_countries)
```

---
As usual, there is a tidyverse alternative to view spatial objects and edit the associated data
```{r spdplyr, echo = TRUE}
library(spdplyr)
pkmng_countries
```

---
With this package, you can use tidyverse functions to manipulate your spatial data:
```{r spdplyr 2, echo = TRUE}
#filter out all countries without any Pokemon sightings, and let's rename the column while we're at it:
pkmng_countries %>%
  filter(pkmn_rc > 0) %>%
  rename(Country = NAME)
```

---
Now let's load our point data:
```{r points, echo = TRUE}
pkmng_points <- read_csv("Data/pkmng_points.csv")
pkmng_points
```

---
We are going to create a shapefile from this table.

```{r points 2, echo = TRUE}
library(sp)
xy <- tibble(longitude = pkmng_points$longitude,
             latitude = pkmng_points$latitude)
pkmng_points <- SpatialPointsDataFrame(coords = xy,
                                      data = pkmng_points)
```

---
# GEOJSON

```{r geojson, echo = TRUE}
library(geojsonio)
library(rmapshaper)
pkmng_countries_json <- geojson_json(pkmng_countries) #this can be VERY time-consuming
```
